Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-07-26 paper/paper.tex
typos in paper, ref in bib

In regression analysis, the traditional approach to this problem employs a $t$-test to check whether the regression coefficients differ from zero \citep{student1908probable}.
%
However, this test will not succeed here, since the estimate $\hat{E}$ is obtained from a double regression, and any employed regularization will add a bias to the diagonal of $\hat{E}$.
%
Instead, we treat the binarization of $\hat{E}$ as a clustering problem: separate the elements in the diagonal into a group of ``small values'', and a group of ``large values''.
%
More specifically, we propose to maximize the ratio of inter-group variance and to minimize the intra-group variance, over all possible splits of the diagonal into $p$ largest values and $d_x-p$ smallest values.
%
Letting $m_0$ and $m_1$ be the average values of the two clusters, $p$ and $d_x-p$ their size, and $v$ the total variance of the sample, we select the split maximizing the Sonquist-Morgan \citep{sonquist_morgan} criterion $\frac{p(d_x-p)}{d_x} \frac{(m_1 - m_0)^2}{v}$.
%
To binarize $\hat{E}$, set to one all the $p$ diagonal entries belonging to the ``large values'' group in the decided split, and setting to zero the rest of the $d_x-p$ diagonal entries.

>>>>

In regression analysis, the traditional approach to this problem employs a $t$-test to check whether the regression coefficients differ from zero \citep{student1908probable}.
%
However, this test will not succeed here, since the estimate $\hat{E}$ is obtained from a double regression, and any employed regularization will add a bias to the diagonal of $\hat{E}$.
%
Instead, we treat the binarization of $\hat{E}$ as a clustering problem: separate the elements in the diagonal into a group of ``small values'', and a group of ``large values''.
%
More specifically, we propose to maximize the ratio of inter-group variance and to minimize the intra-group variance, over all possible splits of the diagonal into $p$ largest values and $d_x-p$ smallest values.
%
Letting $m_0$ and $m_1$ be the average values of the two clusters, $p$ and $d_x-p$ their size, and $v$ the total variance of the sample, we select the split maximizing the Sonquist-Morgan \citep{sonquist_morgan} \citep{kass75} criterion $\frac{p(d_x-p)}{d_x} \frac{(m_1 - m_0)^2}{v}$.
%
To binarize $\hat{E}$, set to one all the $p$ diagonal entries belonging to the ``large values'' group in the decided split, and setting to zero the rest of the $d_x-p$ diagonal entries.

### 2019-07-26 paper/synthetic.tex
typos in paper, ref in bib

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{synthetic.pdf}
    \caption{Results of synthetic experiments. Each dot depicts the average value of a metric for JRR ($x$-axis) against a competitor ($y$-axis) for each experiment configuration averaged over $20$ random seeds. The prediction error on $Y$ within (in) and outside (out) of the training distribution are weakly but consistently larger for all of the models as compared to B2B with and without the Sonquist binarization (E). The AUC, measuring the ability of each model to separate causal from noncausal factors, is consistently larger in B2B as compared to the other models.}
    \label{table:synthetic}
\end{figure}

Table~\ref{table:synthetic} summarizes the results for all experimental conditions, metrics, and methods.
%
B2B is the method obtaining best test errors, lowest false positive and false negative discovery about $E$, and overall the best AUC when it comes to detecting causal influences.

>>>>

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\textwidth]{synthetic.pdf}
    \caption{Results of synthetic experiments. Each dot depicts the average value of a metric for B2B ($x$-axis) against a competitor ($y$-axis) for each experiment configuration averaged over $20$ random seeds. The prediction error on $Y$ within (in) and outside (out) of the training distribution are weakly but consistently larger for all of the models as compared to B2B with and without the Sonquist binarization (E). The AUC, measuring the ability of each model to separate causal from noncausal factors, is consistently larger in B2B as compared to the other models.}
    \label{table:synthetic}
\end{figure}

Figure~\ref{table:synthetic} summarizes the results for all experimental conditions, metrics, and methods.
%
B2B is the method obtaining best test errors, lowest false positive and false negative discovery about $E$, and overall the best AUC when it comes to detecting causal influences.
